---
title: "Project Proposal - Doordash Delivery Time Prediction"
author: "Paris Patel"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(modest)
library(ggplot2)
```

# Background

The dataset taken into consideration, "Delivery Time Prediction" was provided as a take-home assignment by Doordash during one of their interview processes. This data is a subset of the deliveries received by Doordash in early 2015 in a few subset of cities. It contains about 200,000 rows, each corresponding to one unique delivery and has sixteen different variables detailing order time and delivery time and other related factors. 

# Objective

The delivery time prediction is similar to determining ETA (expected time of arrival at destination) on Google or Apple maps and expected time of order delivery on e-commerce websites like Amazon and Ebay. For such software applications, accuracy of time estimation is essential to enhance customer experience. Hence, purpose of conducting this study is to learn what attributes could affect delivery time prediction and what would be the best way to predict it. It will also help to understand what business tradeoffs will be accepted and what will be the no-goes for the final model to be selected.

# About the data

## Data Source

This data has been selected from the Data Projects section of Stratascratch. The details of the data can be found here: <https://platform.stratascratch.com/data-projects/delivery-duration-prediction>

```{r}
delivery_data <- read.csv("/Users/parispatel/Desktop/My Folders/Grad/Fall 22/BUS 235C/Project/datasets/historical_data.csv")
head(delivery_data)
```

## Data Collection

This data was collected by Doordash as a part of its daily operations. It dates back to early 2015 and includes subset of the cities where Doordash operated.

## Variables

```{r}
str(delivery_data)

# convert following into categorical values
# market_id, store_id, store_primary_category, order_protocol


```

```{r}
names(delivery_data)
```

## Organizing Data

```{r}
#reorganizing variables 
delivery_data$market_id = as.factor(delivery_data$market_id)
delivery_data$store_id = as.factor(delivery_data$store_id)
delivery_data$store_primary_category = as.factor(delivery_data$store_primary_category)
delivery_data$order_protocol = as.factor(delivery_data$order_protocol)

delivery_data$created_at = as.POSIXct(delivery_data$created_at, tz = "UTC")
delivery_data$actual_delivery_time = as.POSIXct(delivery_data$actual_delivery_time, tz = "UTC")
```

```{r}
print(head(delivery_data$created_at))
```

```{r}
#adding column to calculate actual time taken for delivery
delivery_data$actual_delivery_duration = as.numeric(difftime(delivery_data$actual_delivery_time, delivery_data$created_at, units = "secs"))

head(delivery_data$actual_delivery_duration)
```

```{r}
summary(delivery_data)
```

```{r}
#adding a column for busy_dasher_ratio to get % of dashers unavailable during typical working hours
delivery_data$busy_dasher_ratio = delivery_data$total_busy_dashers/delivery_data$total_onshift_dashers
head(delivery_data$busy_dasher_ratio)
```

```{r}
#estimated time of order placing plus driving time from store to destination
delivery_data$estimated_non_prep_duration = delivery_data$estimated_order_place_duration + delivery_data$estimated_store_to_consumer_driving_duration
head(delivery_data$estimated_non_prep_duration)
```

## Data Cleanup/Handling missing values

```{r}
#checking for missing values
colSums(is.na(delivery_data))
```

```{r}
#function to identify mode of distribution
mode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}
```

```{r}
#replacing missing values of store_primary_category with most frequent value for a given store_id
store_id_set <- unique(delivery_data$store_id)
length(store_id_set)
head(store_id_set)
store_category_mode <- c()
filtered_NA <- filter(delivery_data, !is.na(delivery_data$store_primary_category))
head(filtered_NA$store_primary_category)
store <-unique(filtered_NA$store_id)
length(store)
```

```{r}
for (id in store) {
  filtered_set <- filter(filtered_NA, store_id == id)
  store_category_mode[id] = as.character(mode(filtered_set$store_primary_category))
}
print(head(store_category_mode))
```

```{r}
  delivery_data$store_primary_category <- as.character(delivery_data$store_primary_category)

for(id in store) {
  delivery_data$store_primary_category[is.na(delivery_data$store_primary_category) & delivery_data$store_id == id] <- store_category_mode[id]
}

# using exclude = NULL to add a factor level for missing values.
delivery_data$store_primary_category <- factor(delivery_data$store_primary_category, exclude = NULL)
```

```{r}
colSums(is.na(delivery_data))
```

# Exploratory Data Analysis

## Categorical Values

```{r}
data.frame(
  "Attributes" = c("# Markets", 
                   "# Stores", 
                   "# Cuisines", 
                   "# Order Protocols"),
  "Total Count" = c(length(unique(delivery_data$market_id)),
                    length(unique(delivery_data$store_id)),
                    length(unique(delivery_data$store_primary_category)),
                    length(unique(delivery_data$order_protocol)))
)
```

```{r}
par(mfrow = c(3,3))
boxplot(delivery_data$total_items, main = "Total items in order")
boxplot(delivery_data$num_distinct_items, main = "Total unique items in order")
boxplot(delivery_data$subtotal, main = "Subtotal of each order")
boxplot(delivery_data$min_item_price, main = "Minimum price of an item in order")
boxplot(delivery_data$max_item_price, main = "Max price of an item in order")
boxplot(delivery_data$estimated_order_place_duration, main = "Estimated duration of placing an order")
boxplot(delivery_data$estimated_store_to_consumer_driving_duration, main = "Estimated driving time")
boxplot(delivery_data$actual_delivery_time, main = "Actual Delivery time")
boxplot(delivery_data$busy_dasher_ratio, main = "Busy-to-available Dasher ratio")
```

```{r}
library(ggplot2)

boxplot(delivery_data$actual_delivery_duration, main = "Actual Delivery Duration")
(ggplot(delivery_data, aes(x=actual_delivery_duration)) + geom_bar(fill="black"))
boxplot(delivery_data$actual_delivery_duration, main = "Actual Delivery Duration", ylim = c(50,20000))
(ggplot(delivery_data, aes(x=actual_delivery_duration)) + geom_bar(fill="black") + xlim(50,15000))
```

```{r}
library(lessR)

PieChart(order_protocol, data = delivery_data, main = "Use of different Order Protocols")
```

## Central Tendencies

```{r}
summary(delivery_data)
```

## Preparing data for Model Building

As observed from boxplot of `actual_delivery_duration`, I will remove the one extreme outlier

```{r}
#creating dummy variables for categorical variable order_protocol
delivery_data$order_protocol <- factor(delivery_data$order_protocol, exclude = NULL)
order_protocol_dummy <- model.matrix(~order_protocol-1, data = delivery_data)
# -1 so that intercept dummy column is not created
order_protocol_dummy <- subset(order_protocol_dummy, select= -c(order_protocolNA))
dim(order_protocol_dummy)
```

```{r}
#creating dummy variables for categorical variable market_id
delivery_data$market_id <- factor(delivery_data$market_id, exclude = NULL)
market_id_dummy <- model.matrix(~market_id-1, data = delivery_data)
market_id_dummy <- subset(market_id_dummy, select= -c(market_idNA))
dim(market_id_dummy)
```

```{r}
#creating dummy variables for categorical variable store_primary_category
store_primary_category_dummy <- model.matrix(~store_primary_category-1, data = delivery_data)
store_primary_category_dummy <- subset(store_primary_category_dummy, select= -c(store_primary_categoryNA))
dim(store_primary_category_dummy)
```

```{r}
# removing original columns of categorical variables
model_data = subset(delivery_data, select = -c(created_at, market_id, store_id, store_primary_category, actual_delivery_time, order_protocol))
```

```{r}
#library(dplyr)
model_data = data.frame(model_data, order_protocol_dummy, market_id_dummy, store_primary_category_dummy)
```

```{r}
dim(model_data)
```

```{r}
summary(model_data)
```

### Infinite and Missing Value Corrections

```{r}
# treating infinite values generated in busy_dasher_ratio
summary(model_data$busy_dasher_ratio)
sum(is.infinite(model_data$busy_dasher_ratio))

#replacing infinite values with nan
model_data$busy_dasher_ratio[is.infinite(model_data$busy_dasher_ratio)] <- NA

summary(model_data$busy_dasher_ratio)
sum(is.infinite(model_data$busy_dasher_ratio))

sum(is.na(model_data))

model_data <- tidyr::drop_na(model_data)
dim(model_data)
sum(is.na(model_data))
```

```{r}
summary(model_data$actual_delivery_duration)
```

### Removing redundancy

```{r}
#Takes too much time for computation, hence ran it only once and saved the output.

library(corrplot)

#corrplot(cor(model_data), type = "lower", tl.cex = 0.2)
```

Removing store_primary_categoryindonesian since its standard deviation is zero.

```{r}
model_data <- subset(model_data, select = -store_primary_categoryindonesian)
dim(model_data)
```

Removing correlated data now. I'll keep a threshold of 0.5 to consider it as highly correlated variables.

```{r}
# getting top n correlations 

get_top_correlations <- function(df, n) {
  abs_corr <- cor(df)
  abs_corr[upper.tri(abs_corr, diag = TRUE)] <- NA
  abs_corr <- reshape2::melt(abs_corr, na.rm = TRUE)
  abs_corr$value <- abs(abs_corr$value)
  abs_corr <- abs_corr[order(-abs_corr$value),]
  return(head(abs_corr,n))
}

```

```{r}
get_top_correlations(model_data,20)
```

`total_busy_dashers` and `total_onshift_dashers` are actually represented by `busy_dashers_ratio`. Additionally, markets are also correlated to each other. `store_primary_categoryindonesian` has zero standard deviation.

```{r}
model_data = subset(delivery_data, select = -c(created_at, market_id, store_id, store_primary_category, actual_delivery_time, order_protocol))

#library(dplyr)
model_data = data.frame(model_data, order_protocol_dummy, store_primary_category_dummy)

model_data = subset(model_data, select = -c(total_onshift_dashers, total_busy_dashers, store_primary_categoryindonesian, estimated_non_prep_duration))
#replacing infinite values with nan

model_data$busy_dasher_ratio[is.infinite(model_data$busy_dasher_ratio)] <- NA

model_data <- tidyr::drop_na(model_data)
dim(model_data)
sum(is.na(model_data))
```

```{r}
get_top_correlations(model_data,20)
```

`order_protocol` are correlated to each other. Hence dropping that too.

```{r}
model_data = subset(delivery_data, select = -c(created_at, market_id, store_id, store_primary_category, actual_delivery_time, order_protocol))

#library(dplyr)
model_data = data.frame(model_data, store_primary_category_dummy)

model_data = subset(model_data, select = -c(total_onshift_dashers, total_busy_dashers, store_primary_categoryindonesian, estimated_non_prep_duration))

#replacing infinite values with nan
model_data$busy_dasher_ratio[is.infinite(model_data$busy_dasher_ratio)] <- NA

model_data <- tidyr::drop_na(model_data)
dim(model_data)
sum(is.na(model_data))
```

```{r}
get_top_correlations(model_data,20)
```

```{r}
model_data$percent_distinct_item_total = model_data$num_distinct_items/model_data$total_items
model_data$avg_price_per_item = model_data$subtotal/model_data$total_items
model_data$price_range_of_items = model_data$max_item_price - model_data$min_item_price
model_data <- subset(model_data, select = -c(num_distinct_items, subtotal, max_item_price, min_item_price))
```

```{r}
get_top_correlations(model_data,20)
```

```{r}
dim(model_data)
```

### Multicollinearity Check

Variance Inflation Factor.

```{r}
library(car)

vif_model <- lm(actual_delivery_duration~., model_data)
sort(vif(vif_model), decreasing = TRUE)

barplot(vif(vif_model), horiz = TRUE, col = "blue", main = "Variance Inflation Factor")
#v1 <- as.data.frame(vif(vif_model))
#filter(v1, vif(vif_model)>15)
```

```{r}
#removing variables with high Variance Inflation Factor
aftervif = data.frame()
drop = TRUE
threshold = 15
while(drop == TRUE) {
  vfit = vif(vif_model)
  
  aftervif = plyr::rbind.fill(aftervif,as.data.frame(t(vfit)))
  if(max(vfit)>threshold) {
    vif_model = update(vif_model,as.formula(paste(".","~",".","-",names(which.max(vfit)))))
  }
  else { drop = FALSE}
}

t_aftervif <- as.data.frame(t(aftervif))
t_aftervif[order(-t_aftervif$V1),]

barplot(vif(vif_model), horiz = TRUE, col = "blue", main = "Variance Inflation Factor")
```

```{r}
model_data <- subset(model_data, select = -c(store_primary_categoryamerican))
dim(model_data)
```

```{r}
library(ggplot2)

boxplot(model_data$actual_delivery_duration, main = "Actual Delivery Duration")
(ggplot(model_data, aes(x=actual_delivery_duration)) + geom_bar(fill="black"))
boxplot(model_data$actual_delivery_duration, main = "Actual Delivery Duration", ylim = c(50,60000))
(ggplot(model_data, aes(x=actual_delivery_duration)) + geom_bar(fill="black") + xlim(50,15000))
```

#### Removing data with actual_delivery_time \> 60000 as outliers

```{r}
max(model_data$actual_delivery_duration)
model_data <- subset(model_data, model_data$actual_delivery_duration < 60000)
dim(model_data)
```

# Feature Selection Using Forward step Selection Method

Using forward step selection method to find set of predictors that best explains the variance in the response.

```{r}
library(leaps)

feat_slct <- regsubsets(actual_delivery_duration~., data = model_data, nvmax = 80, method = "forward", really.big = TRUE)

res.sum <- summary(feat_slct)
```

```{r}
par(mfrow = c(1,2))
plot(feat_slct, scale = "adjr2", main = "Adjusted R-squared")
plot(feat_slct, scale = "r2", main = "R-Squared")
```

```{r}
plot(sqrt(res.sum$rss/nrow(model_data)), type = "l", ylab = "RMSE", main = "RMSE Trend with number of variables")
```

```{r}
mean(sqrt(res.sum$rss/nrow(model_data)))
median(sqrt(res.sum$rss/nrow(model_data)))
min(sqrt(res.sum$rss/nrow(model_data)))
```

```{r}
which.max (res.sum$adjr2)
par(mfrow =c(1,2))
plot(sqrt(res.sum$rss/nrow(model_data)) ,xlab=" Number of Variables ",ylab=" RMSE", type="l")
plot(res.sum$adjr2 ,xlab =" Number of Variables ", ylab=" Adjusted RSq",type="l")   
points (58, res.sum$adjr2[58], col ="red",cex = 1, pch = 20)
```

```{r}
coef(feat_slct, 20)
print("'''''''''")
coef(feat_slct, 40)
print("'''''''''")
coef(feat_slct, 58)
```

```{r}
feat_model_20 <- lm(actual_delivery_duration ~ total_items
                    + total_outstanding_orders
                    + estimated_order_place_duration
                    + estimated_store_to_consumer_driving_duration
                    + busy_dasher_ratio
                    + store_primary_categoryalcohol
                    + store_primary_categorybarbecue
                    + store_primary_categorybrazilian
                    + store_primary_categorycafe
                    + store_primary_categoryjapanese
                    + store_primary_categorymexican
                    + store_primary_categorypersian
                    + store_primary_categorypizza
                    + store_primary_categorysalad
                    + store_primary_categorysandwich
                    + store_primary_categoryseafood
                    + store_primary_categorytapas
                    + store_primary_categoryvietnamese
                    + avg_price_per_item
                    + price_range_of_items,
                    data = model_data )


predicted_data_20 <- data.frame(pred = predict(feat_model_20, model_data), actual = model_data$actual_delivery_duration)

RMSE_20 <- sqrt(mean((predicted_data_20$actual - predicted_data_20$pred)^2))
RMSE_20
```

```{r}
feat_model_40 <- lm(actual_delivery_duration~ total_items 
                 + total_outstanding_orders 
                 + estimated_order_place_duration 
                 + estimated_store_to_consumer_driving_duration 
                 + busy_dasher_ratio 
                 + store_primary_categoryalcohol 
                 + store_primary_categoryargentine
                 + store_primary_categoryasian 
                 + store_primary_categorybarbecue
                 + store_primary_categorybrazilian 
                 + store_primary_categorybreakfast 
                 + store_primary_categorybubble.tea 
                 + store_primary_categoryburger
                 + store_primary_categorycafe 
                 + store_primary_categorychinese
                 + store_primary_categoryconvenience.store        
                 + store_primary_categorydessert
                 + store_primary_categorydim.sum        
                 + store_primary_categoryfilipino
                 + store_primary_categoryfrench        
                 + store_primary_categorygreek
                 + store_primary_categoryhawaiian
                 + store_primary_categoryindian
                 + store_primary_categoryitalian
                 + store_primary_categoryjapanese  
                 + store_primary_categorykorean 
                 + store_primary_categorymediterranean        
                 + store_primary_categorymexican 
                 + store_primary_categorypasta         
                 + store_primary_categorypersian 
                 + store_primary_categorypizza       
                 + store_primary_categorysalad 
                 + store_primary_categorysandwich        
                 + store_primary_categoryseafood
                 + store_primary_categorysouthern       
                 + store_primary_categorytapas       
                 + store_primary_categoryturkish 
                 + store_primary_categoryvietnamese     
                 + avg_price_per_item 
                 + price_range_of_items, 
                 data = model_data )

predicted_data_40 <- data.frame(pred = predict(feat_model_40, model_data), actual = model_data$actual_delivery_duration)

RMSE_40 <- sqrt(mean((predicted_data_40$actual - predicted_data_40$pred)^2))
RMSE_40
```

```{r}
feat_model_58 <- lm(actual_delivery_duration ~ total_items
                    + total_outstanding_orders
                    + estimated_order_place_duration
                    + estimated_store_to_consumer_driving_duration
                    + busy_dasher_ratio
                    + store_primary_categoryafghan
                    + store_primary_categoryafrican
                    + store_primary_categoryalcohol
                    + store_primary_categoryargentine
                    + store_primary_categoryasian
                    + store_primary_categorybarbecue
                    + store_primary_categorybrazilian
                    + store_primary_categorybreakfast
                    + store_primary_categorybubble.tea 
                    + store_primary_categoryburger 
                    + store_primary_categorycafe 
                    + store_primary_categorycajun 
                    + store_primary_categorychinese 
                    + store_primary_categorycomfort.food 
                    + store_primary_categoryconvenience.store
                    + store_primary_categorydessert 
                    + store_primary_categorydim.sum 
                    + store_primary_categoryeuropean 
                    + store_primary_categoryfilipino 
                    + store_primary_categoryfrench 
                    + store_primary_categorygerman 
                    + store_primary_categorygreek 
                    + store_primary_categoryhawaiian 
                    + store_primary_categoryindian 
                    + store_primary_categoryitalian 
                    + store_primary_categoryjapanese 
                    + store_primary_categorykorean
                    + store_primary_categorykosher 
                    + store_primary_categorylatin.american 
                    + store_primary_categorymediterranean 
                    + store_primary_categorymexican 
                    + store_primary_categorypasta 
                    + store_primary_categorypersian 
                    + store_primary_categorypizza 
                    + store_primary_categoryrussian 
                    + store_primary_categorysalad 
                    + store_primary_categorysandwich 
                    + store_primary_categoryseafood 
                    + store_primary_categorysingaporean 
                    + store_primary_categorysmoothie
                    + store_primary_categorysoup 
                    + store_primary_categorysouthern 
                    + store_primary_categoryspanish 
                    + store_primary_categorysteak 
                    + store_primary_categorytapas 
                    + store_primary_categorythai 
                    + store_primary_categoryturkish 
                    + store_primary_categoryvegan 
                    + store_primary_categoryvegetarian 
                    + store_primary_categoryvietnamese 
                    + percent_distinct_item_total 
                    + avg_price_per_item 
                    + price_range_of_items,
                    data = model_data )

predicted_data_58 <- data.frame(pred = predict(feat_model_58, model_data), actual = model_data$actual_delivery_duration)

RMSE_58 <- sqrt(mean((predicted_data_58$actual - predicted_data_58$pred)^2))
RMSE_58
```

```{r}
final_data_20 <- subset(model_data, 
                     select = c(total_items, 
                                total_outstanding_orders,
                                estimated_order_place_duration,
                                estimated_store_to_consumer_driving_duration,
                                busy_dasher_ratio, 
                                store_primary_categoryalcohol,
                                store_primary_categorybarbecue, 
                                store_primary_categorybrazilian, 
                                store_primary_categorycafe, 
                                store_primary_categoryjapanese, 
                                store_primary_categorymexican, 
                                store_primary_categorypersian, 
                                store_primary_categorypizza, 
                                store_primary_categorysalad, 
                                store_primary_categorysandwich, 
                                store_primary_categoryseafood, 
                                store_primary_categorytapas, 
                                store_primary_categoryvietnamese, 
                                avg_price_per_item, 
                                price_range_of_items,
                                actual_delivery_duration
                     ))
```

```{r}
final_data_40 <- subset(model_data, 
                     select = c(
                       total_items,
                       total_outstanding_orders,
                       estimated_order_place_duration,
                       estimated_store_to_consumer_driving_duration,
                       busy_dasher_ratio ,
                       store_primary_categoryalcohol ,
                       store_primary_categoryargentine,
                       store_primary_categoryasian ,
                       store_primary_categorybarbecue,
                       store_primary_categorybrazilian ,
                       store_primary_categorybreakfast,
                       store_primary_categorybubble.tea ,
                       store_primary_categoryburger,
                       store_primary_categorycafe ,
                       store_primary_categorychinese,
                       store_primary_categoryconvenience.store ,
                       store_primary_categorydessert,
                       store_primary_categorydim.sum,
                       store_primary_categoryfilipino,
                       store_primary_categoryfrench   ,
                       store_primary_categorygreek,
                       store_primary_categoryhawaiian,
                       store_primary_categoryindian ,
                       store_primary_categoryitalian,
                       store_primary_categoryjapanese,  
                       store_primary_categorykorean ,
                       store_primary_categorymediterranean,   
                       store_primary_categorymexican ,
                       store_primary_categorypasta ,
                       store_primary_categorypersian, 
                       store_primary_categorypizza  ,
                       store_primary_categorysalad ,
                       store_primary_categorysandwich,   
                       store_primary_categoryseafood,
                       store_primary_categorysouthern,     
                       store_primary_categorytapas ,
                       store_primary_categoryturkish, 
                       store_primary_categoryvietnamese, 
                       avg_price_per_item ,
                       price_range_of_items,
                       actual_delivery_duration
                     ))
dim(final_data)
```

```{r}
final_data_58 <- subset(model_data, 
                     select = c(total_items,
                      total_outstanding_orders,
                      estimated_order_place_duration,
                      estimated_store_to_consumer_driving_duration,
                      busy_dasher_ratio,
                      store_primary_categoryafghan,
                      store_primary_categoryafrican,
                      store_primary_categoryalcohol,
                      store_primary_categoryargentine,
                      store_primary_categoryasian,
                      store_primary_categorybarbecue,
                      store_primary_categorybrazilian,
                      store_primary_categorybreakfast,
                      store_primary_categorybubble.tea, 
                      store_primary_categoryburger, 
                      store_primary_categorycafe, 
                      store_primary_categorycajun, 
                      store_primary_categorychinese, 
                      store_primary_categorycomfort.food, 
                      store_primary_categoryconvenience.store,
                      store_primary_categorydessert ,
                      store_primary_categorydim.sum ,
                      store_primary_categoryeuropean ,
                      store_primary_categoryfilipino ,
                      store_primary_categoryfrench ,
                      store_primary_categorygerman ,
                      store_primary_categorygreek ,
                      store_primary_categoryhawaiian, 
                      store_primary_categoryindian ,
                      store_primary_categoryitalian ,
                      store_primary_categoryjapanese ,
                      store_primary_categorykorean,
                      store_primary_categorykosher ,
                      store_primary_categorylatin.american ,
                      store_primary_categorymediterranean ,
                      store_primary_categorymexican ,
                      store_primary_categorypasta ,
                      store_primary_categorypersian, 
                      store_primary_categorypizza ,
                      store_primary_categoryrussian, 
                      store_primary_categorysalad ,
                      store_primary_categorysandwich, 
                      store_primary_categoryseafood ,
                      store_primary_categorysingaporean, 
                      store_primary_categorysmoothie,
                      store_primary_categorysoup ,
                      store_primary_categorysouthern ,
                      store_primary_categoryspanish ,
                      store_primary_categorysteak ,
                      store_primary_categorytapas ,
                      store_primary_categorythai ,
                      store_primary_categoryturkish, 
                      store_primary_categoryvegan, 
                      store_primary_categoryvegetarian, 
                      store_primary_categoryvietnamese ,
                      percent_distinct_item_total ,
                      avg_price_per_item ,
                      price_range_of_items,
                      actual_delivery_duration
                     ))
```

# Model Building

## Standardization

```{r}
library(rsample)
set.seed(8)
splitset <- initial_split(final_data_20, prop = 0.8)
delivery_train_20 = training(splitset)
delivery_test_20 = testing(splitset)
dim(delivery_train_20)
dim(delivery_test_20)
```

```{r}
x_bar_20 <- mean(delivery_train_20[,"actual_delivery_duration"])
sigma_x_20 <- sd(delivery_train_20[,"actual_delivery_duration"])

for (i in 1:(ncol(delivery_test_20)-1)) {
  delivery_test_20[,i] = scale(delivery_test_20[,i], center = mean(delivery_train_20[,i]), scale = sd(delivery_train_20[,i]))
}
head(delivery_test_20)
```

```{r}
# scaling all the variables to bring them into similar range and making interpretation easy

for (i in 1:(ncol(delivery_train_20))) {
    delivery_train_20[, i] = scale(delivery_train_20[, i])
}
head(delivery_train_20)
```

```{r}
library(rsample)
set.seed(8)
splitset <- initial_split(final_data_40, prop = 0.8)
delivery_train_40 = training(splitset)
delivery_test_40 = testing(splitset)
dim(delivery_train_40)
dim(delivery_test_40)
```

```{r}
x_bar_40 <- mean(delivery_train_40[,"actual_delivery_duration"])
sigma_x_40 <- sd(delivery_train_40[,"actual_delivery_duration"])

for (i in 1:(ncol(delivery_test_40)-1)) {
  delivery_test_40[,i] = scale(delivery_test_40[,i], center = mean(delivery_train_40[,i]), scale = sd(delivery_train_40[,i]))
}
head(delivery_test_40)
```

```{r}
# scaling all the variables to bring them into similar range and making interpretation easy

for (i in 1:(ncol(delivery_train_40))) {
    delivery_train_40[, i] = scale(delivery_train_40[, i])
}
head(delivery_train_40)
```

```{r}
library(rsample)
set.seed(8)
splitset <- initial_split(final_data_58, prop = 0.8)
delivery_train_58 = training(splitset)
delivery_test_58 = testing(splitset)
dim(delivery_train_58)
dim(delivery_test_58)
```

```{r}
x_bar_58 <- mean(delivery_train_58[,"actual_delivery_duration"])
sigma_x_58 <- sd(delivery_train_58[,"actual_delivery_duration"])

for (i in 1:(ncol(delivery_test_58)-1)) {
  delivery_test_58[,i] = scale(delivery_test_58[,i], center = mean(delivery_train_58[,i]), scale = sd(delivery_train_58[,i]))
}
head(delivery_test_58)
```

```{r}
# scaling all the variables to bring them into similar range and making interpretation easy

for (i in 1:(ncol(delivery_train_58))) {
    delivery_train_58[, i] = scale(delivery_train_58[, i])
}
head(delivery_train_58)
```

## Linear Regression

```{r}
linear_model_20 <- lm(actual_delivery_duration ~., 
                   data = delivery_train_20 )
summary(linear_model_20)
```

```{r}
linear_model_40 <- lm(actual_delivery_duration ~., 
                   data = delivery_train_40 )
summary(linear_model_40)
```

```{r}
linear_model_58 <- lm(actual_delivery_duration ~., 
                   data = delivery_train_58 )
summary(linear_model_58)
```

```{r}
par(mfrow = c(2,2))
plot(linear_model_20)
```

```{r}
par(mfrow = c(2,2))
plot(linear_model_40)
```

```{r}
par(mfrow = c(2,2))
plot(linear_model_58)
```

## Random Forest

```{r}
library(ranger)
rf_model_20 <- ranger(x = delivery_train_20[-c(ncol(delivery_train_20))],
                   y = delivery_train_20$actual_delivery_duration,
                   seed = 123,
                   num.tree = 500,
                   write.forest = TRUE,
                   replace = TRUE,
                   importance = "impurity")
```

```{r}
rf_model_20
```

```{r}
res_20 <- delivery_train_20$actual_delivery_duration - rf_model_20$predictions
qqnorm(res_20)
qqline(res_20)
```

```{r}
plot(res_20, pch = 20, cex = 1, main = "Residual vs Fitted")
```

```{r}
rf_model_40 <- ranger(x = delivery_train_40[-c(ncol(delivery_train_40))],
                   y = delivery_train_40$actual_delivery_duration,
                   seed = 123,
                   num.tree = 500,
                   write.forest = TRUE,
                   replace = TRUE,
                   importance = "impurity")
```

```{r}
rf_model_40
```

```{r}
res_40 <- delivery_train_40$actual_delivery_duration - rf_model_40$predictions
qqnorm(res_40)
qqline(res_40)
```

```{r}
plot(res_40, pch = 20, cex = 1, main = "Residual vs Fitted")
```

```{r}
rf_model_58 <- ranger(x = delivery_train_58[-c(ncol(delivery_train_58))],
                   y = delivery_train_58$actual_delivery_duration,
                   seed = 123,
                   num.tree = 500,
                   write.forest = TRUE,
                   replace = TRUE,
                   importance = "impurity")
```

```{r}
rf_model_58
```

```{r}
res_58 <- delivery_train_58$actual_delivery_duration - rf_model_58$predictions
qqnorm(res_58)
qqline(res_58)
```

```{r}
plot(res_58, pch = 20, cex = 1, main = "Residual vs Fitted")
```

## Gradient Boosting

Tried 1000 trees first, and there was no improvement after 550 trees. hence seelct 500 trees for final model.

```{r}
library(gbm)
set.seed(123)
gb_model_20 <- gbm(actual_delivery_duration ~.,
                data = delivery_train_20,
                distribution = "gaussian",
                n.trees = 500,
                shrinkage = 0.01)
summary(gb_model_20)
```

```{r}
head(sqrt(gb_model_20$train.error))
sqrt(mean(gb_model_20$train.error))
head(gb_model_20$fit)
```

```{r}
plot(sqrt(gb_model_20$train.error), ylab = "RMSE for each step", main = "RMSE Plot")
```

```{r}
res_gb_20 <- delivery_train_20$actual_delivery_duration - predict(gb_model_20, delivery_train_20)
qqnorm(res_gb_20)
qqline(res_gb_20)
```

```{r}
plot(res_gb_20, pch = 20, cex = 1, main = "Residual vs Fitted")
```

```{r}
set.seed(43)
gb_model_40 <- gbm.fit(x = delivery_train_40[-c(ncol(delivery_train_40))],
                y = delivery_train_40$actual_delivery_duration,
                distribution = "gaussian",
                n.trees = 500,
                shrinkage = 0.01)
summary(gb_model_40)
```

```{r}
head(sqrt(gb_model_40$train.error))
sqrt(mean(gb_model_40$train.error))
head(gb_model_40$fit)
```

```{r}
plot(sqrt(gb_model_40$train.error), ylab = "RMSE for each step", main = "RMSE Plot")
```

```{r}
res_gb_40 <- delivery_train_40$actual_delivery_duration - predict(gb_model_40, delivery_train_40)
qqnorm(res_gb_40)
qqline(res_gb_40)
```

```{r}
plot(res_gb_40, pch = 20, cex = 1, main = "Residual vs Fitted")
```

```{r}
set.seed(25)
gb_model_58 <- gbm.fit(x = delivery_train_58[-c(ncol(delivery_train_58))],
                y = delivery_train_58$actual_delivery_duration,
                distribution = "gaussian",
                n.trees = 500,
                shrinkage = 0.01)
summary(gb_model_58)
```

```{r}
head(sqrt(gb_model_58$train.error))
sqrt(mean(gb_model_58$train.error))
head(gb_model_58$fit)
```

```{r}
plot(sqrt(gb_model_58$train.error), ylab = "RMSE for each step", main = "RMSE Plot")
```

```{r}
res_gb_58 <- delivery_train_58$actual_delivery_duration - predict(gb_model_58, delivery_train_58)
qqnorm(res_gb_58)
qqline(res_gb_58)
```

```{r}
plot(res_gb_58, pch = 20, cex = 1, main = "Residual vs Fitted")
```

## Prediction

### In-sample

```{r}
linear_pred_train_20 <- predict(linear_model_20, delivery_train_20)
rf_pred_train_20 <- predict(rf_model_20, delivery_train_20)$predictions
gb_pred_train_20 <- predict(gb_model_20, delivery_train_20)
```

```{r}
linear_pred_train_40 <- predict(linear_model_40, delivery_train_40)
rf_pred_train_40 <- predict(rf_model_40, delivery_train_40)$predictions
gb_pred_train_40 <- predict(gb_model_40, delivery_train_40)
```

```{r}
linear_pred_train_58 <- predict(linear_model_58, delivery_train_58)
rf_pred_train_58 <- predict(rf_model_58, delivery_train_58)$predictions
gb_pred_train_58 <- predict(gb_model_58, delivery_train_58)
```

### Unscaling

```{r}
unscaling <- function(vector,xbar,sigma_x) {
  return((vector * sigma_x) + x_bar)
}
```

``` {r}
delivery_train_20$actual_delivery_duration <- unscaling(delivery_train_20$actual_delivery_duration,x_bar_20, sigma_x_20)
delivery_train_40$actual_delivery_duration <- unscaling(delivery_train_40$actual_delivery_duration,x_bar_40, sigma_x_40)
delivery_train_58$actual_delivery_duration <- unscaling(delivery_train_58$actual_delivery_duration,x_bar_58, sigma_x_58)
```

``` {r}
linear_pred_train_20 <- unscaling(linear_pred_train_20,x_bar_20, sigma_x_20)
rf_pred_train_20 <- unscaling(rf_pred_train_20,x_bar_20, sigma_x_20)
gb_pred_train_20 <- unscaling(gb_pred_train_20,x_bar_20, sigma_x_20)

linear_pred_train_40 <- unscaling(linear_pred_train_40,x_bar_40, sigma_x_40)
rf_pred_train_40 <- unscaling(rf_pred_train_40,x_bar_40, sigma_x_40)
gb_pred_train_40 <- unscaling(gb_pred_train_40,x_bar_40, sigma_x_40)

linear_pred_train_58 <- unscaling(linear_pred_train_58,x_bar_58, sigma_x_58)
rf_pred_train_58 <- unscaling(rf_pred_train_58,x_bar_58, sigma_x_58)
gb_pred_train_58 <- unscaling(gb_pred_train_58,x_bar_58, sigma_x_58)

head(delivery_train_20$actual_delivery_duration)
```

```{r}

data.frame(
  No_of_Variables = c(20,40,58),
  Linear.RMSE = c(sqrt(mean((delivery_train_20$actual_delivery_duration - linear_pred_train_20)^2)),
                  sqrt(mean((delivery_train_40$actual_delivery_duration - linear_pred_train_40)^2)),
                  sqrt(mean((delivery_train_58$actual_delivery_duration - linear_pred_train_58)^2))),
  Random_forest.RMSE = c(sqrt(mean((delivery_train_20$actual_delivery_duration - rf_pred_train_20)^2)),
                         sqrt(mean((delivery_train_40$actual_delivery_duration - rf_pred_train_40)^2)),
                         sqrt(mean((delivery_train_58$actual_delivery_duration - rf_pred_train_58)^2))),
  Gradient_boost.RMSE = c(sqrt(mean((delivery_train_20$actual_delivery_duration - gb_pred_train_20)^2)),
                          sqrt(mean((delivery_train_40$actual_delivery_duration - gb_pred_train_40)^2)),
                          sqrt(mean((delivery_train_58$actual_delivery_duration - gb_pred_train_58)^2)))
)
```

### Out-of-sample

```{r}
linear_pred_20 <- predict(linear_model_20, delivery_test_20)
rf_pred_20 <- predict(rf_model_20, delivery_test_20)$predictions
gb_pred_20 <- predict(gb_model_20, delivery_test_20)
```

```{r}
linear_pred_40 <- predict(linear_model_40, delivery_test_40)
rf_pred_40 <- predict(rf_model_40, delivery_test_40)$predictions
gb_pred_40 <- predict(gb_model_40, delivery_test_40)
```

```{r}
linear_pred_58 <- predict(linear_model_58, delivery_test_58)
rf_pred_58 <- predict(rf_model_58, delivery_test_58)$predictions
gb_pred_58 <- predict(gb_model_58, delivery_test_58)
```

```{r}
head(linear_pred_20)
print("*************************************************************")
head(rf_pred_20)
print("*************************************************************")
head(gb_pred_20)
print("*************************************************************")

head(linear_pred_40)
print("*************************************************************")
head(rf_pred_40)
print("*************************************************************")
head(gb_pred_40)
print("*************************************************************")

head(linear_pred_58)
print("*************************************************************")
head(rf_pred_58)
print("*************************************************************")
head(gb_pred_58)
```

```{r}
linear_pred_20 <- unscaling(linear_pred_20,x_bar_20, sigma_x_20)
rf_pred_20 <- unscaling(rf_pred_20,x_bar_20, sigma_x_20)
gb_pred_20 <- unscaling(gb_pred_20,x_bar_20, sigma_x_20)

linear_pred_40 <- unscaling(linear_pred_40,x_bar_40, sigma_x_40)
rf_pred_40 <- unscaling(rf_pred_40,x_bar_40, sigma_x_40)
gb_pred_40 <- unscaling(gb_pred_40,x_bar_40, sigma_x_40)

linear_pred_58 <- unscaling(linear_pred_58,x_bar_58, sigma_x_58)
rf_pred_58 <- unscaling(rf_pred_58,x_bar_58, sigma_x_58)
gb_pred_58 <- unscaling(gb_pred_58,x_bar_58, sigma_x_58)
```

```{r}
head(linear_pred_20)
head(rf_pred_20)
head(gb_pred_20)
head(linear_pred_40)
head(rf_pred_40)
head(gb_pred_40)
head(linear_pred_58)
head(rf_pred_58)
head(gb_pred_58)
```

## RMSE

```{r}
data.frame(
  No_of_Variables = c(20,40,58),
  Linear.RMSE = c(sqrt(mean((delivery_train_20$actual_delivery_duration - linear_pred_20)^2)),
                  sqrt(mean((delivery_train_40$actual_delivery_duration - linear_pred_40)^2)),
                  sqrt(mean((delivery_train_58$actual_delivery_duration - linear_pred_58)^2))),
  Random_forest.RMSE = c(sqrt(mean((delivery_train_20$actual_delivery_duration - rf_pred_20)^2)),
                         sqrt(mean((delivery_train_40$actual_delivery_duration - rf_pred_40)^2)),
                         sqrt(mean((delivery_train_58$actual_delivery_duration - rf_pred_58)^2))),
  Gradient_boost.RMSE = c(sqrt(mean((delivery_train_20$actual_delivery_duration - gb_pred_20)^2)),
                          sqrt(mean((delivery_train_40$actual_delivery_duration - gb_pred_40)^2)),
                          sqrt(mean((delivery_train_58$actual_delivery_duration - gb_pred_58)^2)))
)
```

<https://doordash.engineering/2021/04/28/improving-eta-prediction-accuracy-for-long-tail-events/>
